---
title: "correlator"
author: "Andrew Ba Tran"
date: "July 27, 2015"
output: html_document
---

An [analysis](http://trendct.github.io/ct-tickets/correlator.html) for the TrendCT quiz: [Quiz: What correlates with traffic tickets per town?](http://trendct.org/2015/08/04/tickets-correlation-quiz/)

The traffic ticket data is from the [Connecticut Racial Profiling Prohibition Project](http://ctrp3.ctdata.org/rawdata/) hosted at the [CT Data Collaborative](http://ctdata.org/) which collects traffic incident reports between October 2013 and September 2014.

The [vehicle density](https://github.com/trendct/ct-tickets/data/blob/master/dmvt.csv) and [traffic count](https://github.com/trendct/ct-tickets/data/blob/master/traffic_count.csv) data is from the Department of Transportation.

[Population](https://github.com/trendct/ctnamecleaner), [town area](https://github.com/trendct/ct-tickets/data/blob/master/area.csv), [commuter count](https://github.com/trendct/ct-tickets/data/blob/master/commuters.csv), and [median income](https://github.com/trendct/ct-tickets/data/blob/master/medianincome.csv) data is from the US Census.

The business revenue and taxes [data](https://github.com/trendct/ct-tickets/data/blob/master/salestax.csv) is from the [Connecticut Data Portal](https://data.ct.gov/Government/Sales-and-Use-Tax-per-Town-by-NAICS-2013-and-2014-/rkm7-uwkb).

The Dunkin' Donuts [data](https://github.com/trendct/ct-tickets/data/blob/master/dunks.csv) was scraped from the Dunkin' Donuts website.

Just interested in the big joined Tickets dataframe? Here's the [data](https://github.com/trendct/ct-tickets/data/blob/master/big_tickets_dataframe.csv).

```{r}
library(stringr)
library(dplyr)
library(ctnamecleaner)
library(tidyr)
library(ggplot2)
library(DT)
```

Bringing in the data, cleaning it, setting it up for analysis

```{r}
#Bringing in the traffic count statistics from the Department of Transportation
traffic <- read.csv("data/traffic_count.csv", stringsAsFactors=FALSE)

traffic_towns <- tapply(traffic$ADT, traffic$Town_Name, mean)

towns <- data.frame(traffic_towns)

towns$town <- row.names(towns)
towns <- towns[c("town", "traffic_towns")]
colnames(towns) <- c("town", "ADT Average")
roads <- data.frame(table(traffic$Town_Name))
colnames(roads) <- c("town", "roads")

towns <- left_join(towns, roads)
colnames(towns) <- c("town", "ADT", "roads")
towns$ADT <- round(towns$ADT, digits=2)
types <- data.frame(table(traffic$Town_Name, traffic$FClass))
colnames(types) <- c("town", "ftype", "number")
types$ftype <- as.character(types$ftype)
types$ftype <- gsub("1", "f1", types$ftype)
types$ftype <- gsub("2", "f2", types$ftype)
types$ftype <- gsub("3", "f3", types$ftype)
types$ftype <- gsub("4", "f4", types$ftype)
types$ftype <- gsub("5", "f5", types$ftype)
types$ftype <- gsub("6", "f6", types$ftype)
types$ftype <- gsub("7", "f7", types$ftype)
types <- spread(types, ftype, number)

towns <- left_join(towns, types)
towns$town <- str_trim(towns$town)
dmvt <- read.csv("data/dmvt.csv", stringsAsFactors=FALSE)
dmvt$town <- str_to_title(dmvt$Town)
dmvt$DVMT <- gsub(",", "", dmvt$DVMT)
dmvt$DVMT <- as.numeric(dmvt$DVMT)
dmvt <- dmvt[,-1]
dmvt <- dmvt[,-1]

towns <- left_join(towns, dmvt)

#

  # Bringing in square miles per town
  area <- read.csv("data/area.csv", stringsAsFactors=FALSE)
colnames(area) <- c("town", "sqmiles")
area$town <- str_trim(area$town)
towns <- left_join(towns, area)

# Bringing in population stats
towns <- ctpopulator(town, towns)
towns$density <- towns$pop2013/towns$sqmiles

#

# Bring in retail sales by town
sales <- read.csv("data/salestax.csv", stringsAsFactors=FALSE)
salestx <- subset(sales, NAICS.Industry.Code=="All NAICS Codes")
sales2014 <- subset(salestx, Calendar.Year==2014)
retail <- tapply(sales2014$Retail.Sales.of.Goods, sales2014$Municipality, sum)
write.csv(retail, "data/temp_retail.csv")
retail <- read.csv("data/temp_retail.csv", stringsAsFactors=FALSE)
file.remove("data/temp_retail.csv")
colnames(retail) <- c("town", "retail.tax")
retail <- retail[-1,]
retail$town <- str_trim(retail$town)
towns <- left_join(towns, retail)


#retail$town <- rownames(retail)
#colnames(retail) <- c("retail.sales", "town")
#retail <- retail[c("town", "retail.sales")]
#towns2 <- left_join(retail, towns)
#towns <- left_join(retail, towns)


taxes <- tapply(sales2014$Total.Tax.Due.at.6.35., sales2014$Municipality, sum)
write.csv(taxes, "data/temp_taxes.csv")
taxes <- read.csv("data/temp_taxes.csv", stringsAsFactors=FALSE)
file.remove("data/temp_taxes.csv")
colnames(taxes) <- c("town", "total.tax")
taxes <- taxes[-1,]
taxes$town <- str_trim(taxes$town)

towns <- left_join(towns, taxes)

# Bringing in commuters stats from the US Census
commuters <- read.csv("data/commuters.csv", stringsAsFactors=FALSE)

commuters_filtered <- commuters[c("name", "B08006001...Total.", "B08006002...Car..truck..or.van.")]
commuters_filtered <- commuters_filtered[-1,]
commuters_filtered <- commuters_filtered[-1,]

colnames(commuters_filtered) <- c("town","commute_total", "drivers")
commuters_filtered$town <- sub(" town,.*", "", commuters_filtered$town)
commuters_filtered$percent.drivers <- round((commuters_filtered$drivers/commuters_filtered$commute_total)*100, digits=2)
commuters_filtered$town <- str_to_upper(commuters_filtered$town)

towns <- left_join(towns, commuters_filtered)

towns$percent.drivers.all <- round((towns$drivers/towns$pop2013)*100, digits=2)

  # Bringing in median income data 
  income <- read.csv("data/medianincome.csv", stringsAsFactors=FALSE)
income <- income[,-1]
income <- income[-1,]
colnames(income) <- c("town", "median.income", "error")
income <- income[c("town", "median.income")]
income$town <- sub(" town,.*", "", income$town)
income$town <- str_to_upper(income$town)

towns <- left_join(towns, income)


# Bringing in the tickets
incidents <- read.csv("http://ctrp3viz.s3.amazonaws.com/data/Connecticut_r1.csv", stringsAsFactors=FALSE)
warnings <- data.frame(table(incidents$Department.Name, incidents$InterventionDispositionCode))
colnames(warnings) <- c("Department", "Disposition", "Freq")
warnings_df <- warnings %>% spread(Disposition, Freq)
warnings_df$total <- warnings_df$I + warnings_df$M + warnings_df$N + warnings_df$U + warnings_df$V + warnings_df$W
warnings_df$warnings <- warnings_df$V + warnings_df$W
warnings_df$tickets <- warnings_df$total - warnings_df$warnings

warnings_df$perc_warnings <- round((warnings_df$warnings/warnings_df$total)*100, digits=2)
warnings_df$perc_tickets <- 100 - warnings_df$perc_warnings
warnings_df_table <- warnings_df[c("Department", "total", "tickets", "warnings", "perc_tickets", "perc_warnings")]

#exclude Stamford because of unreliable data

warnings_df_table <- filter(warnings_df_table, Department!="Stamford")

tix <- warnings_df_table

colnames(tix) <- c("town", "stops", "tickets", "warnings", "perc_tickets", "perc_warnings")
tix$town <- str_to_upper(tix$town)

tix_df <- left_join(towns, tix)

tix_df <- na.omit(tix_df)

# Bringin police department staffing data
pers <- read.csv("data/police_dept.csv", stringsAsFactors=FALSE)
pers$sworn <- pers$sworn.male+ pers$sworn.female
incidents_by_dept <- data.frame(table(incidents$Department.Name))
colnames(incidents_by_dept) <- c("name", "incidents")
incidents_by_dept$name <- toupper(incidents_by_dept$name)

incidents_by_dept <- left_join(incidents_by_dept, pers)

incidents_by_dept$inc.rate <- round((incidents_by_dept$incidents/incidents_by_dept$sworn), digits=2)

dept_staff <- incidents_by_dept[c("name", "sworn.male", "sworn.female", "sworn", "total")]
colnames(dept_staff) <- c("town", "sworn.male", "sworn.female", "sworn", "dept.total")

tix_df <- left_join(tix_df, dept_staff)

tix_only <- tix_df[,25]
tix_save <- tix_df

colnames(tix_df) <- c("town", "Average daily traffic", "Number of roads", 
                      "Number of interstate roads", "Number of arterials: Freeways and expressways",
                      "Number of arterials: Other", "Minor arterial roads", "Major collector roads",
                      "Minor collector roads", "Local roads", "Daily vehicle miles traveled", 
                      "Miles of road", "Mileage", "Town square miles", "Town population", 
                      "Population per square mile", "Retail tax 2014", "Total tax 2014",
                      "Total commuters", "Total drivers", "Drivers as a percent of commuters", 
                      "Drivers as a percent of total population", "Town median income", 
                      "Total traffic stops in 2014", "Total tickets issued in 2014", "Total warnings issued in 2014", 
                      "Percent of tickets given out of all stops", 
                      "Percent of warnings given out of all stops", "Sworn male officers", 
                      "Sworn female officers", "Total sworn officers", "Total police department employees")

write.csv(tix_df, "data/big_tickets_dataframe.csv")
datatable(tix_df)
```

### Generating a list of correlations to tickets by town
```{r}
tix_list <- 2:ncol(tix_df)
for (i in tix_list) {
  row_name <- colnames(tix_df[i])
  corre <- cor(tix_only, tix_df[,i], use="pairwise.complete.obs")
  loop_array <- data.frame(row_name, corre)
  if (i == 2) {
    array <- loop_array } else 
    { array <- rbind(array, loop_array) }
}
write.csv(array, "data/tickets_correlation_list.csv")
datatable(array)
```

### Plot it out. Comparing tickets

```{r}
for (i in tix_list) {
title <- paste("Tickets vs ", array$row_name[i-1], ": ", round(array$corre[i-1], digits=2))
p <- ggplot(tix_df, aes(tix_df[,25], tix_df[,i])) + geom_point() + ggtitle(title) + ylab(array$row_name[i-1]) + xlab("Tickets")
plot(p)
}
```

### Comparing warnings

```{r}
warn_only <- tix_df[,26]
for (i in tix_list) {
  row_name <- colnames(tix_df[i])
  corre <- cor(warn_only, tix_df[,i], use="pairwise.complete.obs")
  loop_array <- data.frame(row_name, corre)
  if (i == 2) {
    array <- loop_array } else 
    { array <- rbind(array, loop_array) }
}

for (i in tix_list) {
  title <- paste("Warnings vs ", array$row_name[i-1], ": ", round(array$corre[i-1], digits=2))
  p <- ggplot(tix_df, aes(tix_df[,26], tix_df[,i])) + geom_point() + ggtitle(title) + ylab(array$row_name[i-1]) + xlab("Warnings")
plot(p)
}
```

### Comparing traffic stops total

```{r}
stops_only <- tix_df[,24]
for (i in tix_list) {
  row_name <- colnames(tix_df[i])
  corre <- cor(stops_only, tix_df[,i], use="pairwise.complete.obs")
  loop_array <- data.frame(row_name, corre)
  if (i == 2) {
    array <- loop_array } else 
    { array <- rbind(array, loop_array) }
}

for (i in tix_list) {
  title <- paste("Traffic stops vs ", array$row_name[i-1], ": ", round(array$corre[i-1], digits=2))
  p <- ggplot(tix_df, aes(tix_df[,24], tix_df[,i])) + geom_point() + ggtitle(title) + ylab(array$row_name[i-1]) + xlab("Traffic stops")
plot(p)
}
```

### Dunkin Donuts


```{r}

dunks <- read.csv("data/dunks.csv", stringsAsFactors=FALSE)
dunks <- ctnamecleaner(Towns, dunks, case="upper")
colnames(dunks) <- c("d.town", "dunkinrate","dunks", "town")
dunks$town <- str_to_upper(dunks$town)
dunkin <- left_join(tix_df, dunks)

dunkin <- na.omit(dunkin)

dcorr <- cor(dunkin$dunks, dunkin[,25])

  title <- paste("Traffic stops vs Dunkin' Donuts: ", round(dcorr, digits=2))
  p <- ggplot(dunkin, aes(dunkin[,25], dunkin$dunks)) + geom_point() + ggtitle(title) + ylab("Dunkin' Donut stores") + xlab("Tickets")
plot(p)
```

### Correlograms

```{r}
library(corrgram)
txd1 <- tix_save[c("tickets", "ADT", "roads", "f1", "f2", "f3", "f4", "f5", "f6", "f7", "DVMT", "Lane.Miles", "Mileage")]

corrgram(txd1, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Tickets and roads")

txd2 <- tix_save[c("tickets", "sqmiles", "pop2013", "density", "retail.tax", "total.tax", "commute_total", "drivers", "percent.drivers", "percent.drivers.all", "median.income")]

corrgram(txd2, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Tickets and taxes and drivers and income")

txd3 <- tix_save[c("tickets", "stops", "warnings", "perc_tickets", "perc_warnings", "sworn.male", "sworn.female", "sworn", "dept.total")]
  
corrgram(txd3, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Tickets and warnings and stops and staffing")
```