---
title: "Traffic tickets in Connecticut overview"
author: "Andrew Ba Tran"
date: "July 3, 2015"
output: html_document
---

An analysis for the TrendCT story: [When do police most often ticket drivers? And for what?](http://trendct.org/2015/07/06/police-tickets-in-connecticut/).

The data is from the [Connecticut Racial Profiling Prohibition Project](http://ctrp3.ctdata.org/rawdata/) hosted at the [CT Data Collaborative](http://ctdata.org/) which collects traffic incident reports between October 2013 and September 2014.

```{r, echo=FALSE}
library(lubridate)
library(ggplot2)
library(knitr)
```


```{r}
# Bringing the data in
incidents <- read.csv("incidents_slim.csv", stringsAsFactors=FALSE)

# This is a slightly smaller version of the original version
# If the original raw data is what you prefer, use this command instead of the one above
# incidents <- read.csv("http://ctrp3viz.s3.amazonaws.com/data/Connecticut_r1.csv", stringsAsFactors=FALSE)
```

## Most popular ticket type

```{r}
c <- ggplot(incidents, aes(x=StatutoryReasonForStop))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c + theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Cleaning up the dates and time
incidents$RealDate <- as.POSIXct(as.Date(incidents$InterventionDateTime, origin="1899-12-30"))
incidents$RealTime <- format(as.POSIXct((incidents$InterventionTime) * 86400, origin = "1970-01-01"), "%H:%M")
incidents$RealTime <- hm(incidents$RealTime)
incidents$Day.of.Week <- as.factor(incidents$Day.of.Week)
levels(incidents$Day.of.Week) <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
```

## Which month has the most tickets?

```{r}
incidents$month <- month(incidents$RealDate, label=TRUE)

c <- ggplot(incidents, aes(x=month))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c
```

## Which day of the week gets the most tickets?

```{r}
c <- ggplot(incidents, aes(x=Day.of.Week))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c
```

## Which specific day of the year gets the most tickets? 

```{r}
incidents$calendar <- floor_date(incidents$RealDate, "day")

cal_days <- data.frame(table(incidents$calendar))
colnames(cal_days) <- c("Date", "Tickets")
library(dygraphs)
library(xts)

# This is to convert the time series into another format called eXtensible Time Series
tix <- xts(cal_days[,-1],order.by=as.POSIXct(cal_days$Date))
colnames(tix) <- "Tickets"
dygraph(tix, main="Tickets by day in Connecticut") %>%
  dyOptions(stepPlot = TRUE, fillGraph=TRUE) %>%
  dyRangeSelector()
```

## 10 days that got the most tickets

```{r}
kable(head(cal_days[order(-cal_days$Tickets),],10))

#Most day of the month gets the most tickets?
incidents$day_num <- day(incidents$RealDate)
c <- ggplot(incidents, aes(x=day_num))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c

# That's not very useful-- 11 months of the year have 30 days while half of that has 31.
# Let's try to be more accurate

# Let's focus on the last week, first week, and everything in between as middle weeks. 
# What's the average ticket per day based on those categories?

incidents$cal_month <- floor_date(incidents$RealDate, "month")
cal_months <- data.frame(table(incidents$cal_month))

incidents$cal_week <- floor_date(incidents$RealDate, "week")

cal_week <- data.frame(table(incidents$cal_week))

incidents$cal_days <- days_in_month(incidents$RealDate)

incidents$day_num <- day(incidents$RealDate)

day_num <- data.frame(sort(table(incidents$day_num), descending=TRUE))

longmonth <- subset(incidents, cal_days==31)
midmonth <- subset(incidents, cal_days==30)
shortmonth <- subset(incidents, cal_days==28)

##31 DAY MONTH ANALYSIS HERE
longmonth_end <- subset(longmonth, (day_num==25 | day_num==26 | day_num==27 | 
                                      day_num==28 | day_num==29 | day_num==30 | day_num==31))

last_seven_long <- nrow(longmonth_end)/7

longmonth_start <- subset(longmonth, (day_num==1 | day_num==2 | day_num==3 | 
                                      day_num==4 | day_num==5 | day_num==6 | day_num==7))

first_seven_long <- nrow(longmonth_start)/7

longmonth_mid <- subset(longmonth, (day_num!=1 & day_num!=2 & day_num!=3 & 
                                        day_num!=4 & day_num!=5 & day_num!=6 & day_num!=7 &
                                      day_num!=25 & day_num!=26 & day_num!=27 & 
                                      day_num!=28 & day_num!=29 & day_num!=30 & day_num!=31))

mid_seven_long <- nrow(longmonth_mid)/17
```

### Average number of tickets per day in a 31 day month

```{r}
avg_long <- data.frame(first_seven_long, mid_seven_long, last_seven_long)
colnames(avg_long) <- c("First week", "Middle weeks", "Last week")
kable(avg_long)
```



```{r}
##30 DAY MONTH ANALYSIS HERE
midmonth_end <- subset(midmonth, (day_num==24 | day_num==25 | day_num==26 | 
                                      day_num==27 | day_num==28 | day_num==29 | day_num==30))

last_seven_mid <- nrow(midmonth_end)/7

midmonth_start <- subset(midmonth, (day_num==1 | day_num==2 | day_num==3 | 
                                        day_num==4 | day_num==5 | day_num==6 | day_num==7))

first_seven_mid <- nrow(midmonth_start)/7

midmonth_mid <- subset(midmonth, (day_num!=1 & day_num!=2 & day_num!=3 & 
                                      day_num!=4 & day_num!=5 & day_num!=6 & day_num!=7 &
                                      day_num!=24 & day_num!=25 & day_num!=26 & 
                                      day_num!=27 & day_num!=28 & day_num!=29 & day_num!=30))

mid_seven_mid <- nrow(midmonth_mid)/16

```

### Average number of tickets per day in a 30 day month

```{r}
avg_mid <- data.frame(first_seven_mid, mid_seven_mid, last_seven_mid)
colnames(avg_mid) <- c("First week", "Middle weeks", "Last week")
kable(avg_mid)
```



```{r}
##FEBRUARY ANALYSIS HERE
shortmonth_end <- subset(shortmonth, (day_num==22 | day_num==23 | day_num==24 | 
                                    day_num==25 | day_num==26 | day_num==27 | day_num==28))

last_seven_short <- nrow(shortmonth_end)/7

shortmonth_start <- subset(shortmonth, (day_num==1 | day_num==2 | day_num==3 | 
                                      day_num==4 | day_num==5 | day_num==6 | day_num==7))

first_seven_short <- nrow(shortmonth_start)/7

shortmonth_mid <- subset(shortmonth, (day_num!=1 & day_num!=2 & day_num!=3 & 
                                    day_num!=4 & day_num!=5 & day_num!=6 & day_num!=7 &
                                    day_num!=22 & day_num!=23 & day_num!=24 & 
                                    day_num!=25 & day_num!=26 & day_num!=27 & day_num!=28))

mid_seven_short <- nrow(shortmonth_mid)/14
```

### Average number of tickets per day in February

```{r}
avg_short <- data.frame(first_seven_short, mid_seven_short, last_seven_short)
colnames(avg_short) <- c("First week", "Middle weeks", "Last week")
kable(avg_short)
```


### Average (of the averages) number of tickets per day based on time of month
```{r}
avg_last <- (last_seven_short+last_seven_mid+last_seven_long)/3
avg_mid <- (mid_seven_short+mid_seven_mid+mid_seven_long)/3
avg_start<- (first_seven_short+first_seven_mid+first_seven_long)/3

avg <- data.frame(avg_start, avg_mid, avg_last)
colnames(avg) <- c("First week", "Middle weeks", "Last weeks")
kable(avg)
```

### Overall, it looks like the daily average of tickets increases through the month.
#### Except in months with 30 days of the year. Then the trend reverses.


