---
title: "warnings_analysis"
author: "Andrew Ba Tran"
date: "July 15, 2015"
output: html_document
---

An [analysis](http://trendct.github.io/ct-tickets/warnings_analysis.html) for the TrendCT story: [Who gets off with a warning after a traffic stop in Connecticut?](http://trendct.org/2015/07/16/who-gets-off-with-a-warning-after-a-traffic-stop-in-connecticut)

The data is from the [Connecticut Racial Profiling Prohibition Project](http://ctrp3.ctdata.org/rawdata/) hosted at the [CT Data Collaborative](http://ctdata.org/) which collects traffic incident reports between October 2013 and September 2014.

```{r}
library(lubridate)
library(ggplot2)
library(dplyr)
library(DT)
library(knitr)

incidents <- read.csv("http://ctrp3viz.s3.amazonaws.com/data/Connecticut_r1.csv", stringsAsFactors=FALSE)

warnings <- incidents[grepl("Warn",incidents$InterventionDispositionReasonText),]

# Hm, only 18 departments. Let's take a look at them as a whole

the18 <- subset(incidents, Department.Name=="Brookfield" | Department.Name=="CAPITOL POLICE" | 
                Department.Name=="Coventry" | Department.Name=="Derby" | Department.Name == "East Hampton" |
                Department.Name=="Easton" | Department.Name=="Middlebury" | Department.Name == "Naugatuck" |
                Department.Name=="Plymouth" | Department.Name=="Redding" | Department.Name == "SCSU" |
                Department.Name=="Seymour" | Department.Name=="Suffield" | Department.Name == "Thomaston" |
                Department.Name=="Weston" | Department.Name=="Windsor Locks" | Department.Name == "Winsted" |
                Department.Name=="Wolcott")

kable(head(the18[,1:6]))
```

**Observations**

* There were `r nrow(the18)` traffic stops in this data set

***

## OK, let's split up the dataset to those with warnings and those without


```{r}

the18_non <- the18[!grepl("Warn",the18$InterventionDispositionReasonText),]
the18_inc <- data.frame(table(the18$StatutoryReasonForStop))
colnames(the18_inc) <- c("citation", "total")

the18_warn <- data.frame(table(warnings$StatutoryReasonForStop))
colnames(the18_warn) <- c("citation", "warnings")
the18_df <- left_join(the18_inc, the18_warn)

the18_df$percent_warning <- round((the18_df$warnings/the18_df$total)*100, digits=2)

datatable(the18_df)
```

**Observations**

* Police were more lenient on drivers who were pulled over for defective lights, display of plates, and window tint
* Police were most harsh on drivers pulled over for a suspended license. 
* Only 6 percent of those pulled over got away with a warning
* Driving without a license, without wearing a seatbelt or while using a cell phone also lead to many more tickets than warnings. 
* Registration violations also prompt a high proportion of tickets to warnings.

***

## Stats on each department that gave out warnings

```{r}

department <- data.frame(table(the18$Department.Name))
colnames(department) <- c("Department", "Total.Incidents")
dept_non <- data.frame(table(the18_non$Department.Name))
colnames(dept_non) <- c("Department", "Total.Tickets")
dept_warn <- data.frame(table(warnings$Department.Name))
colnames(dept_warn) <- c("Department", "Total.Warnings")
department <- left_join(department, dept_non)
department <- left_join(department, dept_warn)
department$percent.tickets <- round((department$Total.Tickets/department$Total.Incidents)*100, digits=2)
department$percent.warnings <- round((department$Total.Warnings/department$Total.Incidents)*100, digits=2)
write.csv(department, "department_warnings.csv")
datatable(department)
```

**Observations**

* Suffield gives out the most warnings of any town. Nearly 90 percent of all traffic stops lead to a warning
* Middlebury is also nearly as generous
* Windsor Locks, Coventry, and Thomaston are not nearly as nice. More than 60 percent of their stops end with tickets.

***

## Interesting. Ok let's look at age

```{r}
warnings$SubjectAge <- as.numeric(warnings$SubjectAge)
the18_non$SubjectAge <- as.numeric(the18_non$SubjectAge)
forkable <- data.frame(table(warnings$SubjectAge))
colnames(forkable) <- c("Age", "Tickets")
datatable(forkable)
```

### Oh gross. The ages are pretty messed up. Have to take out the weird ones

```{r}
warningsb <- subset(warnings, SubjectAge > 0 & SubjectAge < 100)
# That takes out about 600 from warnings
the18_nonb <- subset(the18_non, SubjectAge > 0 & SubjectAge < 100)
# That takes out about 60 for ticketed

```

* Average age for drivers getting warnings is `r mean(warningsb$SubjectAge)` 
* Average age for drivers getting tickets is `r mean(the18_nonb$SubjectAge)` 

***

## Plotting out warnings by age

```{r warn, fig.width=8, fig.height=5}
c <- ggplot(warningsb, aes(x=SubjectAge))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c
```

## Plotting out tickets by age

```{r tix, fig.width=8, fig.height=5}
c <- ggplot(the18_nonb, aes(x=SubjectAge))
c <- c + geom_histogram(colour="darkred",fill="white", binwidth=1)
c

```

## Does age change between men and women for tickets or warnings?

```{r charts, fig.width=8, fig.height=5}
# All tickets
gender18 <- data.frame(table(the18_nonb$SubjectAge,the18_nonb$SubjectSexCode))
colnames(gender18) <- c("age", "gender", "tickets")
ggplot(gender18, aes(age, y=tickets, group=gender, colour=gender)) +
  ggtitle("Tickets by gender total") +
  geom_line() +
  geom_point()

ggplot(gender18, aes(age, tickets, group=gender, fill=gender)) + ggtitle("Tickets by gender as a percent") + geom_area(position="fill")

# Warnings
genderw <- data.frame(table(warningsb$SubjectAge,warningsb$SubjectSexCode))
colnames(genderw) <- c("age", "gender", "tickets")
ggplot(genderw, aes(age, y=tickets, group=gender, colour=gender)) +
  ggtitle("Warnings by gender total") +
  geom_line() +
  geom_point()

ggplot(genderw, aes(age, tickets, group=gender, fill=gender)) + ggtitle("Warnings by gender as a percent") + geom_area(position="fill") 

```

## OK, age of driver doesn't mean much

```{r}
all_gender <- data.frame(table(the18_non$SubjectSexCode))
warning_gender <- data.frame(table(warnings$SubjectSexCode))

all_gender$perc <- round((all_gender$Freq/sum(all_gender$Freq))*100, digits=2)
warning_gender$perc <- round((warning_gender$Freq/sum(warning_gender$Freq))*100, digits=2)
kable(warning_gender)

```

## What about race? 

```{r}
#Have to refer to the data dictionary here...
all_race$perc <- round((all_race$Freq/sum(all_race$Freq))*100, digits=2)
the18$Race <- paste(the18$SubjectRaceCode, the18$SubjectEthnicityCode)

index <- c("A H", "A M", "A N", "B H", "B M", "B N", 
           "I H", "I M", "I N", "W H", "W M", "W N")

values <- c("Hispanic", "Middle Eastern", "Asian", "Hispanic", "Middle Eastern", "Black", 
            "Hispanic", "Middle Eastern", "Indian", "Hispanic", "Middle Eastern", "White")

# for all
the18$Def_Race <- values[match(the18$Race, index)]
all_race <- data.frame(table(the18$Def_Race))
colnames(all_race) <- c("race", "incidents")
all_race$percent_incidents <- (all_race$incidents/sum(all_race$incidents))*100
all_race$percent_incidents <- round(all_race$percent_incidents, digits=2)

# for all minus warnings
the18_non$Race <- paste(the18_non$SubjectRaceCode, the18_non$SubjectEthnicityCode)
the18_non$Def_Race <- values[match(the18_non$Race, index)]
all_race_non <- data.frame(table(the18_non$Def_Race))
colnames(all_race_non) <- c("race", "tickets")
all_race_non$percent_tickets <- (all_race_non$tickets/sum(all_race_non$tickets))*100
all_race_non$percent_tickets <- round(all_race_non$percent_tickets, digits=2)

# now just for those who got warnings
warnings$Race <- paste(warnings$SubjectRaceCode, warnings$SubjectEthnicityCode)
warnings$Def_Race <- values[match(warnings$Race, index)]
all_race_warn <- data.frame(table(warnings$Def_Race))
colnames(all_race_warn) <- c("race", "warnings")
all_race_warn$percent_warnings <- (all_race_warn$warnings/sum(all_race_warn$warnings))*100
all_race_warn$percent_warnings <- round(all_race_warn$percent_warnings, digits=2)

all <- left_join(all_race, all_race_non)
all <- left_join(all, all_race_warn)
write.csv(all, "race_warnings.csv")
kable(all)

```
