---
title: "warnings_analysis"
author: "Andrew Ba Tran"
date: "July 15, 2015"
output: html_document
---

An [analysis](http://trendct.github.io/ct-tickets/warnings_analysis.html) for the TrendCT story: [Who gets off with a warning after a traffic stop in Connecticut?](http://trendct.org/2015/07/16/who-gets-let-off-with-a-warning-after-a-traffic-stop-by-police-in-connecticut/)

The data is from the [Connecticut Racial Profiling Prohibition Project](http://ctrp3.ctdata.org/rawdata/) hosted at the [CT Data Collaborative](http://ctdata.org/) which collects traffic incident reports between October 2013 and September 2014.

```{r}
library(lubridate)
library(ggplot2)
library(dplyr)
library(DT)
library(knitr)
library(stringr)
library(tidyr)
incidents <- read.csv("http://ctrp3viz.s3.amazonaws.com/data/Connecticut_r1.csv", stringsAsFactors=FALSE)

#warnings <- incidents[grepl("Warn",incidents$InterventionDispositionReasonText),]

```

**Observations**

* There were `r nrow(incidents)` traffic stops in this data set

***

## OK, let's take a look at departments and tickets vs warnings


```{r}
incidents$InterventionDispositionCode <- str_trim(incidents$InterventionDispositionCode)

# Graph 1. Tickets versus warnings
index <- c("I", "M", "N", "U", "V", "W")

values <- c("Ticket", "Ticket", "Ticket", "Ticket", "Warning", "Warning")

# for all
incidents$What <- values[match(incidents$InterventionDispositionCode, index)]

violations <- data.frame(table(incidents$StatutoryReasonForStop, incidents$What))
colnames(violations) <- c("violation", "type", "freq")
violations_df <- violations %>% spread(type, freq)
violations_df$perc_ticket <- round((violations_df$Ticket/(violations_df$Ticket+violations_df$Warning))*100, digits=2)
violations_df$perc_warning <- round((violations_df$Warning/(violations_df$Ticket+violations_df$Warning))*100, digits=2)
datatable(violations_df)

```

**Observations**

* Police were more lenient on drivers who were pulled over for defective lights, display of plates, and equipment violations.
* Police were most harsh on drivers pulled over for a suspended license. 
* Only 9 percent of those pulled over for suspended license got away with a warning
* Driving without registration, without wearing a seatbelt or while using a cell phone also lead to many more tickets than warnings. 

***

## Stats on each department that gave out warnings

```{r}

warnings <- data.frame(table(incidents$Department.Name, incidents$InterventionDispositionCode))
colnames(warnings) <- c("Department", "Disposition", "Freq")
warnings_df <- warnings %>% spread(Disposition, Freq)
warnings_df$total <- warnings_df$I + warnings_df$M + warnings_df$N + warnings_df$U + warnings_df$V + warnings_df$W
warnings_df$warnings <- warnings_df$V + warnings_df$W
warnings_df$tickets <- warnings_df$total - warnings_df$warnings

warnings_df$perc_warnings <- round((warnings_df$warnings/warnings_df$total)*100, digits=2)
warnings_df$perc_tickets <- 100 - warnings_df$perc_warnings
warnings_df_table <- warnings_df[c("Department", "total", "tickets", "warnings", "perc_tickets", "perc_warnings")]
write.csv(warnings_df_table, "data/department_warnings.csv")
datatable(warnings_df_table)
```

**Observations**

* Out of all stops, ECSU's most likely to give out warnings than tickets. 
* Putnam and Middlebury are towns with police that give out the most warnings-- nearly 9 out of 10 stops lead to a verbal or written warning.
* Suffield and Portland
* Danbury, Meriden, and Hartford are not nearly as nice. More than 60 percent of their stops end with tickets.

***

## Interesting. Ok let's look at age

```{r age, fig.width=9, fig.height=5}
# by age
incidents_all <- incidents
incidents$SubjectAge <- as.numeric(incidents$SubjectAge)

# We have to take out the outlier ages, first. There are some weird negative ages and such

incidents <- subset(incidents, SubjectAge > 14 & SubjectAge < 100)

incidents_age <- data.frame(table(incidents$SubjectAge,incidents$What))
colnames(incidents_age) <- c("age", "type", "freq")

incidents_age_df <- incidents_age %>% spread(type, freq)
incidents_age_df$perc_ticket <- round((incidents_age_df$Ticket/(incidents_age_df$Ticket+incidents_age_df$Warning))*100, digits=2)
incidents_age_df$perc_warning <- round((incidents_age_df$Warning/(incidents_age_df$Ticket+incidents_age_df$Warning))*100, digits=2)

warnings_only <- subset(incidents, What=="Warning")
tickets_only <- subset(incidents, What=="Ticket")

ggplot(incidents_age, aes(age, freq, group=type, fill=type)) + ggtitle("Tickets and warnings by percent") + geom_area(position="fill")

ggplot(incidents_age, aes(age, y=freq, group=type, colour=type)) +
  ggtitle("Warnings by age total") +
  geom_line() +
  geom_point()

```

## Does age change between men and women for warnings?

```{r charts, fig.width=9, fig.height=5}

warnings_gender <- data.frame(table(warnings_only$SubjectAge,warnings_only$SubjectSexCode))
colnames(warnings_gender) <- c("age", "gender", "freq")

ggplot(warnings_gender, aes(age, freq, group=gender, fill=gender)) + ggtitle("Warnings by gender(percent)") + geom_area(position="fill")

warnings_gender_df <- warnings_gender %>% spread(gender, freq)
warnings_gender_df$perc_male <- round((warnings_gender_df$M/(warnings_gender_df$M+warnings_gender_df$F))*100, digits=2)
warnings_gender_df$perc_women <- round((warnings_gender_df$F/(warnings_gender_df$M+warnings_gender_df$F))*100, digits=2)

ggplot(warnings_gender, aes(age, y=freq, group=gender, colour=gender)) +
  ggtitle("Warnings by age total") +
  geom_line() +
  geom_point()
```

## OK, what about tickets by gender?

```{r gender, fig.width=9, fig.height=5}

tickets_gender <- data.frame(table(tickets_only$SubjectAge,tickets_only$SubjectSexCode))
colnames(tickets_gender) <- c("age", "gender", "freq")

tickets_gender_df <- tickets_gender %>% spread(gender, freq)
tickets_gender_df$perc_male <- round((tickets_gender_df$M/(tickets_gender_df$M+tickets_gender_df$F))*100, digits=2)
tickets_gender_df$perc_women <- round((tickets_gender_df$F/(tickets_gender_df$M+tickets_gender_df$F))*100, digits=2)

ggplot(tickets_gender, aes(age, freq, group=gender, fill=gender)) + ggtitle("Tickets by gender(percent)") + geom_area(position="fill")

ggplot(tickets_gender, aes(age, y=freq, group=gender, colour=gender)) +
  ggtitle("Tickets by age total") +
  geom_line() +
  geom_point()

```

```{r agegender, fig.width=9, fig.height=5}
tickets_gender <- data.frame(table(tickets_only$SubjectAge,tickets_only$SubjectSexCode))
colnames(tickets_gender) <- c("age", "gender", "freq")

tickets_gender_df <- tickets_gender %>% spread(gender, freq)
tickets_gender_df$perc_male <- round((tickets_gender_df$M/(tickets_gender_df$M+tickets_gender_df$F))*100, digits=2)
tickets_gender_df$perc_women <- round((tickets_gender_df$F/(tickets_gender_df$M+tickets_gender_df$F))*100, digits=2)

ggplot(tickets_gender, aes(age, freq, group=gender, fill=gender)) + ggtitle("Tickets by gender(percent)") + geom_area(position="fill")

ggplot(tickets_gender, aes(age, y=freq, group=gender, colour=gender)) +
  ggtitle("Tickets by age total") +
  geom_line() +
  geom_point()

```

## What about race? 

```{r}

incidents$Race <- paste(incidents$SubjectRaceCode, incidents$SubjectEthnicityCode)

index_race <- c("A H", "A M", "A N", "B H", "B M", "B N", 
           "I H", "I M", "I N", "W H", "W M", "W N")

values_race <- c("Hispanic", "Middle Eastern", "Asian", "Hispanic", "Middle Eastern", "Black", 
            "Hispanic", "Middle Eastern", "Indian", "Hispanic", "Middle Eastern", "White")

incidents$Def_Race <- values_race[match(incidents$Race, index_race)]

race_df <- data.frame(table(incidents$Def_Race, incidents$What))
colnames(race_df) <- c("Race", "Type", "Freq")
race_df_spread <- race_df %>% spread(Type, Freq)
race_df_spread$perc_ticket <- round((race_df_spread$Ticket/(race_df_spread$Ticket+race_df_spread$Warning))*100, digits=2)
race_df_spread$perc_warning <- round((race_df_spread$Warning/(race_df_spread$Ticket+race_df_spread$Warning))*100, digits=2)
write.csv(race_df_spread, "data/race_warnings.csv")
kable(race_df_spread)
```
